ABC - HTW Fits

### 2023-11-29 Model/_abc_htw.qmd
restructure abc fitting process

select <- dplyr::select; mutate <- dplyr::mutate 
d <- readRDS(here("data/dPrune-01-19-23.rds"))
levels(d$condit)
# Prepare the data for analysis
dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))
ds <- d %>% filter(expMode %in% c("train","train-Nf","test-Nf","test-train-nf")) %>% 
filter(!id %in% unique(dtest$id[dtest$nBand<5])) %>% 
select(id,condit,catOrder,feedbackType,expMode,trial,gt.train,vb,band,bandInt,lowBound,highBound,input,vx,dist,vxb) 
dst <- ds %>% filter(expMode=="train")
dst <- dst %>%
  group_by(id, vb) %>%
  mutate(trial_band = row_number())
```

```{r}
dst <- dst %>% filter(expMode=="train",catOrder=="orig")
vst <- dst %>% filter(condit=="Varied",gt.train<=84) %>% group_by(gt.train,vb,input) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)),
                          trial=gt.train) 
cst <- dst %>% filter(condit=="Constant",gt.train<=84) %>% group_by(gt.train,vb,input) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)),
                          trial=gt.train)

@@

>>>>

select <- dplyr::select; mutate <- dplyr::mutate 
d <- readRDS(here("data/e1_08-21-23.rds"))
levels(d$condit)

dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))

ds <- d %>% filter(expMode %in% c("train","train-Nf","test-Nf","test-train-nf")) |> 
  filter(!id %in% unique(dtest$id[dtest$nBand<5])) |>    
  mutate(input=bandInt) |>
  select(id,condit,bandOrder,expMode,trial,gt.train,vb,band,bandInt,lowBound,highBound,input,vx,dist,vxb)

dst <- ds %>% filter(expMode=="train")
dst <- dst %>%
  group_by(id, vb) %>%
  mutate(trial_band = row_number())
```

```{r}
dst <- dst %>% filter(expMode=="train",bandOrder=="Original")
vst <- dst %>% filter(condit=="Varied",gt.train<=84) %>% group_by(gt.train,vb,input) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)),
                          trial=gt.train) 
cst <- dst %>% filter(condit=="Constant",gt.train<=84) %>% group_by(gt.train,vb,input) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)),
                          trial=gt.train)

@@

n_prior_samples <- 100
prior_samples <- generate_prior(n_prior_samples)

### 2023-11-29 Model/abc_fit.qmd
restructure abc fitting process

tMax=84
avg_dsv <- ds |> filter(condit=="Varied",expMode2=="Train") |> group_by(tr) %>%
  mutate(bandInt2 = sample(rep(c(800, 1000, 1200), each = tMax / 3), tMax, replace = FALSE)[tr]) %>%
  filter(bandInt2 == x) |> select(-bandInt2) |> group_by(tr,condit,x,expMode2) |> summarise(y=mean(y),.groups="keep") |>
  rbind(dsv |> filter(expMode2=="Test") |> group_by(tr,condit,x,expMode2) |> summarise(y=mean(y),.groups="keep") ) |> as.data.table()

avg_dsc <- dsc |> group_by(condit,expMode2,tr,x) |> summarise(y=mean(y),.groups="keep") |> as.data.table()

# ds |> filter(expMode2=="Test") |> group_by(condit,expMode2,tr,x) |> summarise(y=mean(y),n=n(),.groups="keep") |> 
#   ggplot(aes(x=tr,y=y,fill=condit)) + stat_summary(geom="bar",fun=mean,position=position_dodge()) + facet_wrap(~x) + ggtitle("Mean per Trial")

# ds |> filter(expMode2=="Test") |> group_by(condit,expMode2,tr,x) |> summarise(y=median(y),n=n(),.groups="keep") |> 
#   ggplot(aes(x=tr,y=y,fill=condit)) + stat_summary(geom="bar",fun=mean,position=position_dodge()) + facet_wrap(~x) + ggtitle("Median per Trial")

# avg_dsv |> filter(expMode2=="Train") |> group_by(x,tr) |> summarise(y=mean(y),n=n(),.groups="keep") |> 
#   ggplot(aes(x=tr,y=y)) + stat_summary(geom="bar",fun=mean) + facet_wrap(~x) + ggtitle("N per trial")

@@

@@

@@

plan(multisession)
simulated_data <- future_map_dfc(seq_len(nrow(prior_samples)), function(idx) {
  params <- prior_samples[idx, ]
  full_sim_exam(as.data.table(avg_dsv), params$c, params$lr,exam.response, input_layer, output_layer,return_dat = "train_data,test_data")
},.options = furrr_options(seed = T))

dst_obs <- avg_dsv[expMode2=="Test",]$y
dst_obs <- avg_dsv[expMode2=="Test" | expMode2=="Train",]$y

abc_result <- abc(
  target = dst_obs,
  param = prior_samples,
  sumstat =do.call(rbind, simulated_data),
  tol = .03,
  method = "rejection",
  names=colnames(dst_obs)
)

posterior_samples <- abc_result$unadj.values
colnames(posterior_samples) <- c("c", "lr")
head(posterior_samples)

fsv <- full_sim_exam(as.data.table(avg_dsv), c=mode_c, lr=mode_lr,exam.response, input_layer, output_layer,mode="ret")

#fsv <- full_sim(as.data.table(avg_dsv), c=.007, lr=.047,exam.response, input_layer, output_layer)

pgrid <- cbind(prior_samples,dist=abc_result$dist) |> mutate(dist=sqrt(dist))

pgrid |> filter(dist<50) |> ggplot(aes(x=c,y=lr,col=dist)) + geom_point() + viridis::scale_color_viridis()

@@

@@

@@

>>>>

tMax=84
avg_dsv <- ds |> filter(condit=="Varied",expMode2=="Train") |> group_by(tr) %>%
  mutate(bandInt2 = sample(rep(c(800, 1000, 1200), each = tMax / 3), tMax, replace = FALSE)[tr]) %>%
  filter(bandInt2 == x) |> select(-bandInt2) |> group_by(tr,condit,x,expMode2) |> summarise(y=mean(y),.groups="keep") |>
  rbind(dsv |> filter(expMode2=="Test") |> group_by(condit,x,expMode2) |> summarise(y=mean(y),tr=1,.groups="keep") ) |> setDT()

avg_dsc <- ds |> filter(condit=="Constant",expMode2=="Train",tr<=tMax) |> group_by(tr, condit,x,expMode2) |> 
  summarise(y=mean(y),.groups="keep") |> rbind(dsc |> filter(expMode2=="Test") |> group_by(condit,x,expMode2) |> summarise(y=mean(y),tr=1,.groups="keep") ) |> setDT()

@@

abc_loss_plot <- function(abc_result,prior_samples)
{

  pgrid <- cbind(prior_samples,dist=abc_result$dist) |> mutate(dist=sqrt(dist))
  upper_limit <- quantile(pgrid$dist, 0.98)  # 95th percentile
  ggplot(pgrid, aes(x = c, y = lr, color = dist)) +
  geom_point() + 
 #scale_color_brewer(palette = "Spectral",limits = c(NA, upper_limit)) +
  #scale_color_viridis(limits = c(NA, upper_limit)) +
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal()+ ggtitle("Prior Loss")

  post_grid <- cbind(prior_samples[abc_result$region,],dist=abc_result$dist[abc_result$region],k=abc_result$unadj.values)
  ggplot(post_grid, aes(x = c, y = lr, color = dist)) +
  geom_point() + 
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal() + ggtitle("Posterior Loss")

# Define the function
abc_plot <- function(abc_result, data) {

  pgrid <- cbind(prior_samples,dist=abc_v_tetr$dist) |> mutate(dist=sqrt(dist))

  # Extracting posterior samples
  posterior_samples <- abc_result$unadj.values
  colnames(posterior_samples) <- c("c", "lr")
  posterior_samples_long <- tidyr::pivot_longer(as.data.frame(posterior_samples), everything())

  # Generate posterior density plots
  postV <- ggplot(posterior_samples_long, aes(x=value)) +
    geom_density() +
    facet_wrap(~name, scales="free") +
    theme_minimal() +
    labs(x="Value", y="Density", title="Posterior Density Plots")
  print(postV)

  # Compute summary statistics
  sum_stats <- data.frame(
    mean = apply(posterior_samples, 2, mean),
    median = apply(posterior_samples, 2, median),
    mode = apply(posterior_samples, 2, Mode) # Ensure the 'Mode' function is defined
  )
  sum_stats <- tibble::rownames_to_column(sum_stats, var="parameter")
  print(sum_stats)

  # Calculate mode for parameters c and lr
  density_c <- density(posterior_samples[, "c"])
  mode_c <- density_c$x[which.max(density_c$y)]
  density_lr <- density(posterior_samples[, "lr"])
  mode_lr <- density_lr$x[which.max(density_lr$y)]

  # Run simulation with full_sim_exam function
  fsv <- full_sim_exam(as.data.table(data), c=mode_c, lr=mode_lr, exam.response, input_layer, output_layer, mode="ret")

  # Generate plot for simulation results
  plot_fsv <- fsv |> 
    tidyr::pivot_longer(y:pred, names_to="Resp", values_to="vx") |> 
    ggplot(aes(x, vx, fill=Resp, group=Resp)) +
    stat_summary(geom="bar", fun="mean", position=position_dodge()) +
    scale_fill_manual(values=col_themes$wes2) +
    scale_x_continuous(breaks=sort(unique(data$x)), labels=sort(unique(fsv$x))) +
    theme(legend.title = element_blank(), legend.position="top") +
    ggtitle("Fit to Test Only")
  print(plot_fsv)
}

@@

## Varied Average Fits
```{r}

@@

plan(multisession)
sd_v_tetr <- future_map_dfc(seq_len(nrow(prior_samples)), function(idx) {
  params <- prior_samples[idx, ]
  full_sim_exam(as.data.table(avg_dsv), params$c, params$lr,exam.response, input_layer, output_layer,return_dat = "train_data,test_data")
},.options = furrr_options(seed = T))

dst_v_tetr <- avg_dsv[expMode2=="Test" | expMode2=="Train",]$y

abc_v_tetr <- abc(
  target = dst_v_tetr,
  param = prior_samples,
  sumstat =do.call(rbind, sd_v_tetr),
  tol = .01,
  method = "rejection",
  names=colnames(dst_v_tetr)
)

dst_v_te <- avg_dsv[expMode2=="Test",]$y
sd_v_te <- sd_v_tetr[85:90,]

abc_v_te <- abc(
  target = dst_v_te,
  param = prior_samples,
  sumstat =do.call(rbind, sd_v_te),
  tol = .01,
  method = "rejection",
  names=colnames(dst_v_te)
)

abc_plot(abc_v_tetr, avg_dsv) + abc_loss_plot(abc_v_tetr,prior_samples)
abc_plot(abc_v_te, avg_dsv) + abc_loss_plot(abc_v_te,prior_samples)

pgrid <- cbind(prior_samples,dist=abc_v_tetr$dist) |> mutate(dist=sqrt(dist))

pgrid |> filter(dist<80) |> ggplot(aes(x=c,y=lr,col=dist)) + geom_point() + viridis::scale_color_viridis()

@@

str(pgrid)
ggplot(pgrid)+
  geom_tile(aes(x = c, y = lr, fill = dist)) +
  geom_contour(aes(x = alpha, y = beta, z = nLL), color = 'white')

@@

upper_limit <- quantile(pgrid$dist, 0.98)  # 95th percentile
ggplot(pgrid, aes(x = c, y = lr, color = dist)) +
  geom_point() +
 #scale_color_brewer(palette = "Spectral",limits = c(NA, upper_limit)) +
  #scale_color_viridis(limits = c(NA, upper_limit)) +
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal()

ggplot(pgrid |>filter(dist<66)) +
    geom_tile(aes(x = c, y = lr, fill = dist)) +
    geom_contour(aes(x = c, y = lr, z = dist)) +
    scale_fill_viridis() +
    labs(title = "Your Plot Title", x = "c", y = "lr") +
    theme_minimal()

@@

## Constant Average Fits
```{r}

plan(multisession)

sd_c_tetr <- future_map_dfc(seq_len(nrow(prior_samples)), function(idx) {
  params <- prior_samples[idx, ]
  full_sim_exam(as.data.table(avg_dsc), params$c, params$lr,exam.response, input_layer, output_layer,return_dat = "train_data,test_data")
},.options = furrr_options(seed = T))

dst_c_tetr <- avg_dsc[expMode2=="Test" | expMode2=="Train",]$y

abc_c_tetr<- abc(
  target = dst_c_tetr,
  param = prior_samples,
  sumstat =do.call(rbind, sd_c_tetr),
  tol = .01,
  method = "rejection",
  names=colnames(dst_c_tetr)
)

dst_c_te <- avg_dsc[expMode2=="Test",]$y
sd_c_te <- sd_c_tetr[85:90,]

abc_c_te <- abc(
  target = dst_c_te,
  param = prior_samples,
  sumstat =do.call(rbind, sd_c_te),
  tol = .01,
  method = "rejection",
  names=colnames(dst_c_te)
)

abc_plot(abc_c_tetr, avg_dsc)
abc_plot(abc_c_te, avg_dsc)

### 2023-11-29 Model/deOptim_alm.qmd
restructure abc fitting process

purrr::walk(here::here(c("Functions/Display_Functions.R","Functions/misc_model_funs.R",
                         "Functions/alm_core.R","Functions/fit_funs.R")),source)
ds <- readRDS(here::here("data/e1_md_11-06-23.rds")) |> mutate(sbj=id)

>>>>

purrr::walk(here::here(c("Functions/Display_Functions.R","Functions/fun_model.R",
                         "Functions/alm_core.R","Functions/fit_funs.R")),source)
ds <- readRDS(here::here("data/e1_md_11-06-23.rds")) |> mutate(sbj=id)
